### 2011-04-15 Text/paper.tex
Added velocity profiles plus bubble shapes.

In comparison with two-dimensional geometries, the experimental works
\cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
wang-non-circular} show interesting phenomena in three-dimensional geometries. It was found
\cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case. Non axisymmetric case is attributed to the case where the axial radius is
different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary
number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

@@

For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated.
To avoid the mutual influence of bubbles on each other the channel length was chosen as $15
H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a
body force in the framework of the LBM but not with the pressure difference. Therefore, we can not
address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
the gravity effects are negligible and the flow is in the horizontal direction driven by the body
force. 
\begin{figure}[!]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:classical:benchmark}}
\end{figure}[!]
\begin{figure}
\includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The classical benchmark. The semiinfinite gas bubble
propagates through the liquid media. \label{fig:lbm:benchmark}}
\end{figure}

@@

\section{Binary liquid lattice Boltzmann model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implemented as follows:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
relaxation parameter, and $F_i$ is the external force population.

The binary fluid LB model is
based on a free-energy functional \cite{swift,landau}, and operates with two
sets of populations: one to track the pressure and the velocity fields, and another to represent the
phase field $\phi$ indicating gas or liquid.
The equilibrium populations \cite{pooley-contact} are defined as:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k w_i^{\alpha\beta} \partial_{\alpha} \phi\partial_{\beta} \phi, i=1\div Q-1\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), i=1\div Q-1\\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($9$ and
$19$ for the $D2Q9$ and $D3Q19$ models, correspondingly); the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. The particular weights and velocity
sets for the $D2Q9$ and $D3Q15$ models are indicated in Appendix A.

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields.

@@

To further examine the steady-state the velocity in the streamwise direction is plotted. The values
of the velocity are taken on the contour where phase values are taken as $\phi=0$. This corresponds
 to the bubble interface. One can see the contour plot and the corressponing streamwise velocity
component in Fig. \ref{fig:velocity:contour}. The crossection is a plane $x=0$, where $x$ points
towards the reader, $z$ is the streamwise direction. As far as fluxes inside the bubble has
clockwise and counterclockwise directions, see Section \ref{sec:velocity}, one needs to thoroughly
examine only the points corresponding to the front and the rear bubble tips. If one finds it equal
to each other, then it's exactly the velocity with which the bubble propagates in the microchannel. 
\begin{figure}[h!]
\includegraphics[width=\textwidth]{Figures/velocity_interface_contour.eps}\\
\includegraphics[width=\textwidth]{Figures/velocity_interface_values.eps}\\
\caption{The bubble shape and the corresponding streamwise velocities taken on the bubble
interface. Directions $y$ and $z$ are scaled to the half of the channel
$H_{\mathrm{eff}}/2$. One can see that the bubble front and rear tips are propagating with the
same velocities, which is actually the bubble velocity. For more details on velocity patterns
see Section \ref{sec:velocity}. \label{fig:velocity:contour}}
\end{figure}

@@

\section{Non-axisymmetric case}
We first start with the non-axisymmetric case. As it was indicated before for a quite low capillary
number the flow becomes axisymmetric. For
example, according to \cite{heil-threedim} for the capillary
number $Ca=0.01$, the axis and diagonal radiuses are $r_h=0.99$ and $r_d=1.1$ respectively.
Therefore the film thickness in the axial direction is  $0.5\%$ of whole channel width to see the
difference. Our previous calculations \cite{kuzmin-binary2d} show that the film thickness should be
 resolved around $40-50$ percent for calculations to be convergent and grid independent. This
condition certainly implies a large grid with the computational high demand. For instance, to
properly simulate the abovementioned example the grid should be as 
$200x200x3000$ nodes. In this section we want to examine the assumption of the underresolved
interface whether it can produce the consistent film thicknesses for low capillary numbers. The
interface occupies around $1-2$ lattice units.

Here we present simulations for the domain of size $52\mathrm{x}52\mathrm{x}1500$ which simulates
the quarter of the channel using the mirror boundary conditions (See Appendix \ref{append:sym}).
That
results in the effective grid size as $100\mathrm{x}100\mathrm{x}1500$. The simulation is carried
out with the pressure gradient as:
\begin{equation}
\frac{\mathrm{d}P}{\mathrm{d}x}=1.6\,10^{-6}.
\end{equation}
The initial film thickness was taken as $6$ nodes.

One can see the phase profile in the middle of the bubble in
Fig.\ref{fig:quarter:capillary:capillary01}. For comparison reason the axisymmetric phase profile
for $Ca=0.6$ is presented in Fig. \ref{fig:bubble:axisymmetric}. The associated capillary number is
$0.053$ based on
the calculation of the bubble velocity in the center of the bubble. The radiuses in this simulation
are as $R_{axis}=0.93$ and $R_{diag}=1.06$. While $R_{axis}=0.93$ is underestimated parameter and
we can't take into the consideration apriori, one can consider parameter $R_{diag}=1.06$ the
parameters for the capillary number $0.053$ should be according to \citet{heil-threedim} as $0.97$
and $1.05$ according to the VOF simulations of \citet{wang-non-circular}. One can conclude that
even though the axis film thickness is not properly resolved the diagonal values are consistent and
can give an approximate information. 
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/quarter.eps}\\ %\hfill
\includegraphics[width=0.97\textwidth]{Figures/stream_cap01.eps}\\
\caption{Streamlines and phase profile of the quarter of the channel. The capillary number is
$0.053$ and the film thickness is underresolved. \label{fig:quarter:capillary:capillary01}}
\end{figure}
\begin{figure}[ht]
\includegraphics[width=0.97\textwidth]{Figures/square.eps}
\caption{The middle bubble phase profile. One can see that the diagonal radius coincides with the
axial radius. {\color{red} $Ca=0.5$} \label{fig:bubble:axisymmetric}}
\end{figure}

For underresolved schemes we performed a number of simulations on a GPU code with the grid of size
$50\mathrm{x}50\mathrm{750}$. While changing the applied force in a manner indicated in Appendix
\ref{append:scaling} the velocity of the bubble was increased to obtain larger capillary numbers,
see Fig. \ref{fig:underresolved:capillaries}. One can see that the deviation between the diagonal
and axial radiuses are quite significant and they coincide only when the interface is resolved
($Ca\approx 1.0$). In calculations with better grids, i.e. quarter channels simulations on CPU, the
axial radius coincides with the diagonal in the $0.5\%$ range. Moreover, the capillary number curve
resembles already published date of \citet{heil-threedim} in a better way, see Fig.
\ref{fig:capillary:comparison}. 
\begin{figure}[ht]
\includegraphics[width=0.97\textwidth]{Figures/underresolved_capillaries.eps}
\caption{The underresolved profiles axis and diagonal radiuses versus capillary numbers. $50$
percent resolution starts from $Ca=1.0$, which is quite high. One can see it's exactly where
radiuses converge instead of $Ca=0.04$. {\color{red} To delete? However, one can see from
simulations Fig.
\ref{fig:wang:critical:capillary} and Fig.\ref{fig:thickness:variation:ca:ten} the thickness,
especially for Fig.\ref{fig:thickness:variation:ca:ten}, can significantly variate over a bubble
length.} \label{fig:underresolved:capillaries}}
\end{figure}

\section{Velocity pattern}
\label{sec:velocity}
We examined two different velocity patterns to identify change of pattern. Two capillaries numbers
were taken below and above the numbers indicated in Table \ref{table:recirculation:data}, as
$Ca=0.36$ and $Ca=0.88$.Velocity streamlines are
indicated in Fig. \ref{fig:streamlines:pattern}. One can see that the vortex indicated in
Fig.\ref{fig:heil:recirculating} doesn't appear in present calculations. This fact can be
explained because the simulation is carried our for the binary-liquid case but not free-surface
interface presented in the classical Bretherton problem. In the latter case the void phase pushes
the liquid towards the walls. In the present simulation there is no a void phase. Then the
streamlines have the ability to go through the interface and we exactly see this situation occuring.
\begin{figure}[ht]
\includegraphics[width=0.97\textwidth]{Figures/streamlines_ca036.eps}\\
\includegraphics[width=0.97\textwidth]{Figures/streamlines_ca088.eps}
\caption{Streamlines for $Ca=0.36$ and $Ca=0.88$. One can see that there are no vortexes created
before the bubble. It can be explained by the fact that the binary-liquid model is simulated
instead of free-surface interface gas air finger flow.\label{fig:streamlines:pattern}}
\end{figure}

\section{Variation bubble length}
The works of \citet{heil-threedim} of \citet{wang-non-circular} show the variation of the bubble
thickness through the length of
bubble. The length variation is presented in Fig.\ref{fig:thickness:variation:ca:ten} and
Fig.\ref{fig:wang:bubble:shape}. While Fig. \ref{fig:wang:bubble:shape} indicates that the diagonal
radius has a variation for the rather small capillary numbers as $Ca<0.1$, another profile is
presented for the $Ca=10$ and shows the variation for the diagonal radius. We examined a bunch of
different capillaries numbers to present the diagonal radius variations over the length:
The axis and diagonal radiuses variation over the bubble are shown in Figure
\ref{fig:bubble:variation:capillaries}. The rather small capillary numbers one can see the increase
in the rear menisci of the diagonal radius which coincides with the results of
\citet{wang-non-circular}. One can see that variation over the length can be neglected
for relatively high $Ca$ numbers. 
\begin{figure}[ht]
\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_36.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_46.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_61.eps}\hfill
\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_88.eps}\\
\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_one.eps}
\includegraphics[width=0.47\textwidth]{Figures/bubble_length_ca_117.eps}
\caption{The radiuses variations for the capillary number $Ca=0.46$.
One can see the diagonal radius different from the axial radius near the front of the tip which is
somehow coincides with the results of
\citet{heil-threedim}. The results are presented for the
capillary numbers (from left to right, top to bottom)
$0.36$,$0.46$, $0.61$, $0.88$,$1.06$,$1.17$. \label{fig:bubble:variation:capillaries}}
\end{figure}
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/variationoverlength.eps}
\caption{The variation of the bubble length for capillary number $Ca=10$. The x-axis scaled to
half-width of the microchannel height.  \label{fig:thickness:variation:ca:ten}}
\end{figure}

@@

\section{Capillary number}
We performed a number of simulations with quarter CPU code and GPU train simulations and compared
them with data obtained by \cite{heil-threedim}. After performing the simulations we determine the
$R_{axis}$, $R_{diag}$ versus calculated
capillary number
\begin{equation}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}.
\end{equation}
To perform simulations we
started with the following parameters as
$k=0.04$,$A=0.04$,$\Gamma=1.0$. We performed different grids calculations with GPU and CPU codes.
The GPU code (see section \ref{section:bubble:train} about bubble train simulations) simulates a
quarter of the bubble and operates on the grid $100\mathrm{x}100\mathrm{x}1500$. The CPU refined
grid operates on whole $160\mathrm{x}160\mathrm{x}1500$. In this case we see short bubbles as the
length to height ratio is $\frac{500}{160}=3.125$. GPU  For GPU grids due to memory constrains we
chose the grid as $N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$. One can see the
comparison presented in Fig. \ref{fig:capillary:comparison}.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison.eps}
\caption{The comparison between different runs of codes for the axial and the diagonal radiuses
versus capillary numbers. \label{fig:capillary:comparison}}. One can see that the code mimics
behavior of the earlier published results.
\end{figure}

>>>>

In comparison with two-dimensional geometries, the experimental works
\cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
wang-non-circular} show interesting phenomena in three-dimensional geometries. It was found
\cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case. Non axisymmetric case is attributed to the case where the axial radius is
different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[h]
\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary
number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

@@

For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated.
To avoid the mutual influence of bubbles on each other the channel length was chosen as $15
H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a
body force in the framework of the LBM but not with the pressure difference. Therefore, we can not
address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
the gravity effects are negligible and the flow is in the horizontal direction driven by the body
force. 
\begin{figure}[h]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark layout is presented . The semiinfinite gas bubble
propagates through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
free outflow. The bubble propagates with constant velocity $U_{\mathrm{bubble}}$. 
\label{fig:classical:benchmark}}
\end{figure}
\begin{figure}[h]
\includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The lattice Boltzmann benchmark layout is used in present calculcations. The dimensions of
the domain are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}}\times 15 H_{\mathrm{eff}}$. The
length of the bubble was chosen as $5 H_{\mathrm{eff}}$ for the film thickness to stabilize.
Periodic boundary conditions are applied in the streamwise direction. The flow is driven by body
force. \label{fig:lbm:benchmark}}
\end{figure}

@@

\section{Binary liquid lattice Boltzmann model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implemented as follows \cite{ginzburg-boundary-main}:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{propagation step}, 
\end{aligned}
\end{equation}
where $f_i$ is the probability distribution function in the direction $\bm{c_i}$, $\omega$ is the
relaxation parameter, and $F_i$ is the external force population. The external force population
mimics the force behavior, Eq. \ref{full:navier:stokes}.

The binary fluid LB model is
based on a free-energy functional \cite{swift,landau}, and operates with two
sets of populations: one to track the pressure and the velocity fields, and another to represent the
phase field $\phi$ indicating gas or liquid.
The equilibrium populations \cite{pooley-contact} are defined as:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k w_i^{\alpha\beta} \partial_{\alpha} \phi\partial_{\beta} \phi, 1\leq i \leq Q-1\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), \leq i \leq Q-1\\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($9$ and
$19$ for the $D2Q9$ and $D3Q19$ models, correspondingly); the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. The particular weights and velocity
sets for the $D2Q9$ and $D3Q15$ models are indicated in Appendix A.

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\label{full:navier:stokes}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\label{binary:fluid:system}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields.

@@

To further examine the steady-state the velocity in the streamwise direction is plotted. The values
of the velocity are taken on the contour where phase values are taken as $\phi=0$. This corresponds
 to the bubble interface. One can see the contour plot and the corressponing streamwise velocity
component in Fig. \ref{fig:velocity:contour}. The crossection is a plane $x=0$, where $x$ points
towards the reader, $z$ is the streamwise direction. As far as fluxes inside the bubble has
clockwise and counterclockwise directions, see Section \ref{sec:velocity}, one needs to thoroughly
examine only the points corresponding to the front and the rear bubble tips. If one finds it equal
to each other, then it's exactly the velocity with which the bubble propagates in the microchannel. 
\begin{figure}[h!]
\includegraphics[width=\textwidth]{Figures/velocity_interface_contour.eps}\\
\includegraphics[width=\textwidth]{Figures/velocity_interface_values.eps}\\
\caption{The bubble shape and the corresponding streamwise velocities taken on the bubble
interface after $240000$ iterations. Directions $y$ and $z$ are scaled to the half of the channel
$H_{\mathrm{eff}}/2$. One can see that the bubble front and rear tips are propagating with the
same velocities, which is actually the bubble velocity. For more details on velocity patterns
see Section \ref{sec:velocity}. \label{fig:velocity:contour}}
\end{figure}

@@

\section{Velocity pattern}
\label{sec:velocity}
The velocity pattern is known to change its behavior depending on the capillary number. For
relatively low capillary number $Ca<0.6$ there exists a vortex in front of the bubble in the moving
frame moving with the bubble \cite{heil-threedim}. For larger ones it is indicated that there is no
vortex in front of the bubble. The transition capillary number is indicated by
\citet{heil-threedim} and it equals for the rectangular channel $Ca=0.691$. However, the
distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well the
pressure distribution. In the case of the bubble train it is indicated by \citet{kreutzer-taylor}
that the pressure is largely influenced by bubble frequency and slug distance. The present
simulations are conducted for bubble trains and certain differences in terms of changing streamline
patterns are expected. We examined velocity patterns to identify the moment of the streamlines
pattern change. We chose two
representative capillary numbers as $Ca=0.47$ and $Ca=0.63$. 
One can
see a clear transition between associated patterns. The transition capillary number
\cite{heil-threedim} is slightly higher than the simulation results. In present simulations the
transition occurs at $Ca=0.5-0.6$.
\begin{figure}[h]
\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca47.eps}\\
\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca63.eps}\\
\caption{Velocity vector maps for $Ca=0.47$ and $Ca=0.63$. One can see that there are no vortexes
created
before the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
is a different value from $Ca=0.69$ \cite{heil-threedim}. \label{fig:streamlines:pattern}}
\end{figure}

\section{Variation bubble length}
The work of \citet{wang-non-circular} shows the variation of the bubble
thickness through the length of
bubble. The bubble shape in terms of axial and diagonal radii is reconstructed for a number of
capillary numbers, Fig. \ref{fig:bubble:variation:capillaries}. One can see that for smaller
capillary numbers the diagonal radius exhibits a small jump near the rear bubble tip. It coincides
with results of \citet{wang-non-circular}. However, present simulations show smoother behavior of
the jump.  
\begin{figure}[h]
\includegraphics[width=\textwidth]{Figures/bubble_rad_axis.eps}\\
\includegraphics[width=\textwidth]{Figures/bubble_rad_diag.eps}\\
\caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
One can see that the diagonal radius shows the jump near the bubble rear tip. It coincides with the
VOF simulations of \citet{wang-non-circular}. \label{fig:bubble:variation:capillaries}}
\end{figure}

Another interesting phenomenon for the bubble shape was indicated by \citet{heil-threedim}. In
their simulations the non-symmetric shape was observed even for larger capillary number $Ca>4$ near
the bubble front tip. We performed the same simulations.

\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/variationoverlength.eps}
\caption{The variation of the bubble length for capillary number $Ca=10$. The x-axis scaled to
half-width of the microchannel height.  \label{fig:thickness:variation:ca:ten}}
\end{figure}

@@

\section{Capillary number}

While changing the applied force in a manner indicated in Appendix
\ref{append:scaling} the velocity of the bubble was increased to obtain larger capillary numbers,
see Fig. \ref{fig:underresolved:capillaries}. One can see that the deviation between the diagonal
and axial radiuses are quite significant and they coincide only when the interface is resolved
($Ca\approx 1.0$). In calculations with better grids, i.e. quarter channels simulations on CPU, the
axial radius coincides with the diagonal in the $0.5\%$ range. Moreover, the capillary number curve
resembles already published date of \citet{heil-threedim} in a better way, see Fig.
\ref{fig:capillary:comparison}.

We performed a number of simulations with quarter CPU code and GPU train simulations and compared
them with data obtained by \cite{heil-threedim}. After performing the simulations we determine the
$R_{axis}$, $R_{diag}$ versus calculated
capillary number
\begin{equation}
Ca=\frac{\mu_{liq} U_{bubble}}{\gamma}.
\end{equation}
To perform simulations we
started with the following parameters as
$k=0.04$,$A=0.04$,$\Gamma=1.0$. We performed different grids calculations with GPU and CPU codes.
The GPU code (see section \ref{section:bubble:train} about bubble train simulations) simulates a
quarter of the bubble and operates on the grid $100\mathrm{x}100\mathrm{x}1500$. The CPU refined
grid operates on whole $160\mathrm{x}160\mathrm{x}1500$. In this case we see short bubbles as the
length to height ratio is $\frac{500}{160}=3.125$. GPU  For GPU grids due to memory constrains we
chose the grid as $N_x\mathrm{x}N_y\mathrm{x}N_z=52\mathrm{x}52\mathrm{x}750$. One can see the
comparison presented in Fig. \ref{fig:capillary:comparison}.
\begin{figure}
\includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison.eps}
\caption{The comparison between different runs of codes for the axial and the diagonal radiuses
versus capillary numbers. \label{fig:capillary:comparison}}. One can see that the code mimics
behavior of the earlier published results.
\end{figure}
